Management induced changes of soil organic carbon on global croplands

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
final update of figure and text

```{r captionSOCglo, echo=FALSE}
captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time. This makes it hard to pinpoint a specific year for these SOC estimations. In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010."
```

>>>>

```{r captionSOCglo, echo=FALSE}
captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time. This makes it hard to pinpoint a specific year for these SOC estimations. In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010."
```

### 2020-12-14 SOCbudgetPaper/introduction.Rmd
final update of figure and text

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g. growing seasons ([portmann_mirca2000global_2010]) multiple cropping systems ([@waha_multiple_2020]), irrigation systems ([@jagermeyr_water_2015])) or as stylized scenarios (e.g. [@pugh_simulated_2015], [@lutz_simulating_2019]).

>>>>

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g. growing seasons ([@portmann_mirca2000global_2010]) multiple cropping systems ([@waha_multiple_2020]), irrigation systems ([@jagermeyr_water_2015])) or as stylized scenarios (e.g. [@pugh_simulated_2015], [@lutz_simulating_2019]).

### 2020-12-14 SOCbudgetPaper/methods.Rmd
final update of figure and text

  ```{=latex}
    \begin{table*}[h]
    \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
    \tophline
    \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
    \middlehline
    \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                        \cite{faostat_faostat_2016}, \\
                                                                        \cite{schaphoff_lpjml4_2018},\\
                                                                        \cite{weindl_livestock_2017} \end{minipage}} & 
                                                        \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                        LG:C generic values according to Table 5.5B,\\
                                                                        5.5C in \cite{calvo_buendia_ipcc_2019},\\
                                                                        crop-specific N:C from \cite{bodirsky_n2o_2012}
                                                                        \end{minipage}} \\
                              & below-ground residues, &  &  \\
                              & manure                &  &  \\ 
                              \hline
    Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
                              & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
                                  \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
   \bottomhline
   \end{tabular}
   \label{tab:datasourceinputs}
   \belowtable{}
   \end{table*}
```

@@

Manure production and management is based on a revised methodology of [@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modeling. Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets [@weindl_livestock_2017], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see [@calvo_buendia_ipcc_2019]). 
Depending on the feed system we assume manure to be handled in four different ways:
All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing. 
To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields. For manure stored in different animal waste management systems we compiled carbon loss rates depending on the different systems and the associated nitrogen loss rates as specified in [@bodirsky_n2o_2012] (see Sect. \@ref(append:TableclossAWMS) for more details). We limit high application shares at $10\unit{tC\,ha}^{-1}$ to correct for outliers.

>>>>

  ```{=latex}
    \begin{table*}[h]
    \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
    \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
    \tophline
    \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
    \middlehline
    \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                        \cite{faostat_faostat_2016}, \\
                                                                        LPJmL4 \cite{schaphoff_lpjml4_2018},\\
                                                                        \cite{weindl_livestock_2017} \end{minipage}} & 
                                                        \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                        LG:C generic values according to Table 5.5B,\\
                                                                        5.5C in IPCC \cite{calvo_buendia_ipcc_2019},\\
                                                                        crop-specific N:C from \cite{bodirsky_n2o_2012}
                                                                        \end{minipage}} \\
                              & below-ground residues, &  &  \\
                              & manure                &  &  \\ 
                              \hline
    Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
                              & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
                                  IPCC \cite{calvo_buendia_ipcc_2019} and\\ 
                                  CENTURY (\cite{century_model_2000}) \strut \end{minipage}\tabularnewline
   \bottomhline
   \end{tabular}
   \label{tab:datasourceinputs}
   \belowtable{}
   \end{table*}
```

@@

Manure production and management is based on a revised methodology of [@bodirsky_n2o_2012] and is presented here due to its central role in soil carbon modeling. Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets [@weindl_livestock_2017], assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see [@calvo_buendia_ipcc_2019]). 
Depending on the feed system we assume manure to be handled in four different ways:
All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
Whereas for low-income countries, we adopt a share of 25% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing. 
To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields. For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in [@bodirsky_mrcommons_2020] for more details) depending on the different systems and the associated nitrogen loss rates as specified in [@bodirsky_n2o_2012]. We limit high application shares at $10\unit{tC\,ha}^{-1}$ to correct for outliers.

### 2020-12-14 SOCbudgetPaper/results.Rmd
final update of figure and text

Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in [@karstens_socbudgetdata_2020]. In the following, the most important results (see [@karstens_socbudgetpost_2020] for post-processing script) are summarized.

@@

```{r captionSOCmaps, echo=FALSE}
captionSOCmaps <- "(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas. (b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics. Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas. (d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010. "
```

@@

```{r captionSOCscen, echo=FALSE}
captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different scenarios deviating from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage). ConstManagement combines all three modifications. Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state. Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes."
```

@@

Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig. \@ref(fig:SOcinit)). The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901. The $\Delta SOC$ estimation almost halves (from ~26 GtC to ~14 GtC), as does the decrease in $\Delta SOC$ for the period of 1975--2010 (from ~4 GtC to ~2.5 GtC) for the Inital-natveg scenario.

>>>>

Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in [@karstens_model_2020]. In the following, the most important results (see [@karstens_result_2020] for post-processing script) are summarized.

@@

```{r captionSOCmaps, echo=FALSE}
captionSOCmaps <- "(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas. (b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics. Whereas absolute losses $\\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\\mathrm{SCF}$ are often larger in tropical moist areas. (d): $\\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010. "
```

@@

```{r captionSOCscen, echo=FALSE}
captionSOCscen <- "(a) Global $\\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage). ConstManagement combines all three modifications. Note that $\\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state. Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes."
```

@@

Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig. \@ref(fig:SOCinit)). The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901. The $\Delta SOC$ estimation almost halves (from ~26 GtC to ~14 GtC), as does the decrease in $\Delta SOC$ for the period of 1975--2010 (from ~4 GtC to ~2.5 GtC) for the Inital-natveg scenario.
